\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-21 europa-basics.tex
updated constraint table in correct location within europa.tex and corrected arch.tex typos
updates to intro etc

\rx is an adaptive, artificial intelligence based controller and
provides general framework for building reasoning systems for
real-world autonomous vehicles. At MBARI \rx is used for AUV control;
another instantiation of the system is being used for control of a
terrestrial personal robot \cite{pr2, Meeussen:2010dn}. The
development of \rx has been targeted at surveying a number of
oceanographic features which are dynamic and unpredictable
spatio-temporally. Typically this requires our AUVs to balance the
goals of coverage spatially while opportunistically following features
of scientific interest and to do so while being operationally aware of
its own limitations in terms of resources (typically the battery state
of charge) and overall proximity to other observational assets for
obtaining scientific ground-truth. We are therefore interested in
representational frameworks that allow robots to pursue long-term
objectives while choosing short-term gain and can gracefully deal with
exogenous or endogenous change.

To enable this responsiveness to external observations, the agent has
to be able to synthesize plans insitu and to re-plan. We use a
temporal constraint-based planner \eu with a demonstrated legacy of
having flown on NASA space missions \cite{jonsson00,bresina05,
  barreiro09}. Our autonomy architecture brings three key innovations
for AUV adaptation: the use of flexible plan representations,
compositional control with the use of partitioned networks and
off-line learning to inform insitu environmental state estimation. In
this section we detail some of the key concepts relevant to flexible
plan representation. Section \ref{sec:arch} has details the
\rx partitioned architecture and estimation driven vehicle
adaptation.

\paragraph {A brief note on \texttt{EUROPA} history and notation} The
earliest origin of what we will call \eu in this chapter, is from the
\texttt{HSTS} planner \cite{mus94} which was used for a number of
demonstrations for space telescopes as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was the
\texttt{LISP}-based\footnote{This was the first (and only to our
  knowledge) software ever to been written in \texttt{LISP} to be
  flown in space.} planner that was flown as part of the Remote Agent
\texttt{RAX} experiment. Concurrent to \texttt{RAX}, a C++ version of
the planner was implemented at NASA Ames which went on to be used for
the \texttt{MAPGEN} \cite{bresina05} and continues to be used to date,
in the mission-critical uplink process for the command/control of the
Mars Exploration Rovers (MER) mission. A vastly re-factored, higher
performance open source version \cite{europapso} was named \eu and is
the subject of this chapter. Note however, in using the name \eu, we
refer to the \emph{fundamental ideas as well as the planning paradigm}
even if there exists a specific implementation.

\subsection{The Basics}
\label{sec:basics}

\eu uses a \emph{domain model} written in a declarative language
(NDDL: New Domain Description Language), together with initial
conditions and goals also specified in NDDL, to construct a set of
temporal relations that must be true at the start time. These models
include assertions about the physics of the vehicle, i.e how it
responds to external stimulus and internally driven goals. By
propagating these relations forward using Simple Temporal Networks
\cite{dechter91} and applying goal constraints, \eu can select a set
of conditions that should be true in the future, where some of these
conditions will correspond to actions the agent must take. The planner
can backtrack and try another path during search if a goal cannot be
reached while being capable of discarding unachievable goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such variables.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\vskip+0.1cm
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time the most appropriate one for the actual execution
conditions. The fact that constraints are explicitly represented
ensures that through constraint propagation the executive will respect
global limits expressed in the plan (e.g., don't start a task until a
certain condition has been satisfied but still satisfying some global
deadline). Such flexibility is critical when dealing with dynamic
ocean conditions where precise timing of a robotic action might be
indeterminate. Further, the advantage of flexibility can be contrasted
with the consequences of the intrinsic inflexibiliy of traditional
command sequences. Because they are inflexible, sequences must
necessarily be designed considering worst case scenarios.

\eu uses a \emph{state variable} representation to describe the
evolution of state over time. The instantiated history of such state
variable evolution over a temporal horizon we call \emph{timelines}
and which represent a single thread in the execution of a concurrent
system. At any given time each thread can execute a single procedure.
Thus each timeline consists of a sequence of procedures which
encapsulate and describe state evolution; we call these instantiated
atomic entities \emph{tokens}.  A token therefore describes a
procedure invocation, the state variables on which it can occur, the
parameter values of the procedure, and the time values defining the
interval. We allow encapsulation of uncertainty within these tokens
with a range of start and end times and parameters, all of which are
encoded as variables. A constraint solver in turn manipulates these
variables defined in a \eu domain model. For example, consider
a scientific need to take a water samples 100 meters from a hotspot
while an AUV is performing a Yo-Yo in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

% \begin{figure}
% \centering
% \includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
% \caption{\small Temporal relations defined within the planner are
%   based on \texttt{Allen Algebra} relations shown above.}
% \label{fig:allen-algebra}
% \vskip-0.3cm
% \end{figure}

\section{Fundamentals of  Automated Planning}
\label{sec:europa}

@@

% Planning, for our purposes, can be thought of as determining all the
% small tasks that must be carried out in order to accomplish a goal.
To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of the
technique we start with a simple example.

\begin{quotation}

  In the near future, a personal robot sets out to buy a gallon of milk
  This involves a number of tasks: obtain keys, obtain wallet,
  start car, drive to store, find and obtain milk, purchase milk, etc.
  The embedded planner has to have a ``model'' of the world in which it
  lives and has to use the task primitives in this model to structure
  the actions so it achieves its goal. Constraints control when certain
  tasks can or cannot occur. For example the robot must obtain the keys
  and wallet \emph{before} driving to the store and pick up the milk
  \emph{before} purchasing it.

\end{quotation}

@@

\begin{equation}
 Sum(Q_{rc})= \left\{
\begin{array}{l l}
  1 & \forall r \quad \mbox{(only one Queen per row)}\\
  1 & \forall c \quad \mbox{(only one Queen per column)}\\ 
\end{array} \right.
Sum(Q_{r+i,c+i}) = 1\\
Sum(Q_{r-1,c+i}) = 1 \quad \mbox{(only one Queen on each diagonal)}\\
\end {equation}

\comment{fix newline and numbering problem above}

The problem can be solved by finding assignments for all variables
Q$_{rc}$ that satisfy the above constraints. Fig. \ref{fig:nqueens-2} is
a solution found by \eu using that formulation and a specialized search
procedure.

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
AC-3 algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. \comment{A figure of a
  constraint network preferably of the shopping example would be
  useful here. JB to generate a napkin fig}

@@

To generate plans, the CAIP framework uses a plan-space search
approach \cite{ghallab04}, where a partial plan is evolved until all
open goals are achieved, and all inconsistencies are removed, the
specific search algorithm used by \eu will be covered in detail below.

\input{constraints-table}

>>>>

\rx is an adaptive, artificial intelligence based controller and
provides general framework for building reasoning systems for
real-world autonomous vehicles. At MBARI \rx is used for AUV control;
another instantiation of the system is being used for control of a
terrestrial personal robot \cite{pr2, Meeussen:2010dn}. The
development of \rx has been targeted at surveying a number of
oceanographic features which are dynamic and unpredictable
spatio-temporally. Typically this requires our AUVs to balance the
goals of coverage spatially while opportunistically following features
of scientific interest and to do so while being operationally aware of
its own limitations in terms of resources (\eg the battery state of
charge) and overall proximity to other observational assets for
obtaining scientific ground-truth. We are interested in
representational frameworks that allow robots to pursue long-term
objectives while choosing short-term gain and can gracefully deal with
exogenous or endogenous change.

To enable this responsiveness to external observations, the agent has
to be able to synthesize plans insitu and to re-plan. We use the
temporal constraint-based planner \eu with a demonstrated legacy of
having flown on NASA space missions \cite{mus98,jonsson00,bresina05,
  barreiro09}. Our autonomy architecture brings three key innovations
for AUV adaptation: the use of flexible plan representations,
compositional control with the use of partitioned networks and a
demonstrated capability at sea in novel ocean observation methods.

The next few sections will cover \eu in a bottom-up fashion, going
from the generic theoretical frameworks that underlie \eu, to the
specific representation, modeling and algorithms implemented by
it. First, a quick introduction to Constraint Programming and
Constraint-Based Planning is provided.  Next, the specific
representation that \eu uses to perform Constraint-Based Planning is
presented.  After that comes an overview of \eu's architecture and how
it can be embedded in a planning application such as \rx. Finally,
\eu's modeling, inference and search mechanisms are covered in detail
so that the reader can understand how \rx can perform deliberative
planning and react to execution events by embedding and customizing
\eu.

\subsection{Automated Planning and \eu}
\label{sec:planningfound}

@@

To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of the
technique we start with a simple example.

{\small
  \begin{quote}
    In the near future, a personal robot sets out to buy a gallon of
    milk This involves a number of tasks: obtain keys, obtain wallet,
    start car, drive to store, find and obtain milk, purchase milk,
    etc.  The embedded planner has to have a ``model'' of the world in
    which it lives and has to use the task primitives in this model to
    structure the actions so it achieves its goal. Constraints control
    when certain tasks can or cannot occur. For example the robot must
    obtain the keys and wallet \emph{before} driving to the store and
    pick up the milk \emph{before} purchasing it.
\end{quote}

@@

\begin{eqnarray*}
 Sum(Q_{rc}) & = & \left\{
   \begin{array}{l l}
     1 & \forall r \quad \mbox{(only one Queen per row)}\\
     1 & \forall c \quad \mbox{(only one Queen per column)}\\ 
   \end{array} \right. \\
Sum(Q_{r+i,c+i}) & = & 1\\
Sum(Q_{r-i,c-i}) & = & 1 \qquad \mbox{(only one Queen on each diagonal)}\\
\end{eqnarray*}

The problem can be solved by finding assignments for all variables
Q$_{rc}$ that satisfy the above constraints. Fig. \ref{fig:nqueens-2} is
a solution found by \eu using a compact version of that formulation and a specialized search
procedure.

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
\texttt{AC-3} algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. Fig.  \ref{fig:constraintnet} shows a set of constraints on 3 variables
($x$, $y$ and $z$) represented as a constraint network, and their domains before arc consistency is enforced.
After the AC-3 algorithm is executed, the domains for the 3 variables
would be pruned to x=$\{2$\}, y=$\{2$\}, z=$\{2$\}.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint Network}
\label{fig:constraintnet}
\end{figure}

@@

To generate plans, the \texttt{CAIP} framework uses a plan-space search
approach \cite{ghallab04}, where a partial plan is evolved until all
open goals are achieved, and all inconsistencies are removed. The
specific search algorithm used by \eu will be covered in detail below.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "setobook"
%%% End:
